Demes: a standard format for demographic models

### 2022-08-02 paper.tex
Move Example to its own section
Didn't make sense to have this within the Demes section
Remove Parsers appendix
Closes #84

\subsection*{Example: an isolation with migration model}

@@

While the HDM is designed for human readability and
conciseness, the underlying data model suitable for software implementation
(the Machine Data Model, or MDM)
is redundant and exhaustive.
Translation from the HDM to the MDM
requires resolving hierarchically-defined default values and verifying
relationships between populations and the validity of specified parameter
values.
Because this translation and validation requires significant
programming effort, we define a standard software entity as part of the
specification to perform this task (the parser),
which is intended to be shared by programs that support Demes as input.
By maintaining high-quality Demes parsers available as libraries,
we ensure consistency across simulation and inference
software.
Having common parsers also benefits users by providing
consistent and informative error messages for missing values or
issues in formatting.
In addition to a reference implementation written in Python,
we provide high-quality parser implementations in the Python and
C languages (Table~\ref{tab:software}, as discussed in
Appendix~\ref{sec:appendix-parsers}).

@@

\section{Parsers}\label{sec:appendix-parsers}

The implicit value inference and default value propagation required to
fully resolve a Demes HDM description into the corresponding
MDM form is straightforward
to describe, but still requires significant effort to implement in
software. Moreover, there are many constraints on the data model
(for example, epoch \texttt{end\_time} values should be strictly decreasing
within a deme), which must be enforced.
It would not be reasonable to require every program that
takes Demes models as input to implement this logic, as the programming
effort would be significant (limiting adoption)
and it is likely that if there were many implementations some would differ
in their details (harming the software ecosystem).
% GG: this is currently the case for programs accepting ms command lines. E.g.
%  - MaCS interprets -es proportion as 1-proportion compared to ms.
%  - SCRM uses different options for -ema command.
We therefore define
a standard software entity as part of the specification (the parser),
which performs this task and can be shared by programs that
support Demes as an input. By having relatively few high-quality Demes
parsers available as libraries (ideally, one per programming language),
the probability of divergences from the standard is greatly reduced.

The Demes specification precisely defines the required behavior of parsers,
which translate a model described in the HDM into the MDM, where values have been
assigned to all parameters and data
constraints have been checked. The output is formally defined as JSON, but in
practice parsers output an object model that is suitable for the particular
programming language and that can be used directly by the implementing program.
We provide a reference implementation written in Python to resolve any
potential ambiguities and to provide a helpful template for other
implementations, as well as an extensive test suite of examples and the
expected outputs. In addition, we have high-quality parser implementations in
the Python and C languages (published under liberal open source licenses),
providing a solid foundation for the software ecosystem.

>>>>

While the HDM is designed for human readability and
conciseness, the underlying data model suitable for software implementation
(the Machine Data Model, or MDM)
is redundant and exhaustive.
Translation from the HDM to the MDM
requires resolving hierarchically-defined default values and verifying
relationships between populations and the validity of specified parameter
values.
Because this translation and validation requires significant
programming effort, we define a standard software entity as part of the
specification to perform this task (the parser),
which is intended to be shared by programs that support Demes as input.
The Demes specification precisely defines the required behavior of parsers,
and we provide a reference implementation written in Python to resolve any
potential ambiguities, as well as an extensive test suite of examples and the
expected outputs. In addition, we have high-quality parser implementations in
the Python and C languages (Table~\ref{tab:software})
providing a solid foundation for the software ecosystem.
By maintaining high-quality Demes parsers available as libraries,
we ensure consistency across simulation and inference
software.
Having common parsers also benefits users by providing
consistent and informative error messages for missing values or
issues in formatting.

\section*{Example: an isolation with migration model}

@@

@@
